---
title: "Introduction to METEOR"
author: "METEOR Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to METEOR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# METEOR: Meta-analysis Tool for Exploring antimicrobial resistance Outcomes across Realms

## Introduction

Antimicrobial resistance (AMR) is a global health threat that requires coordinated monitoring and analysis across different domains: human, animal, and environment. METEOR (Meta-analysis Tool for Exploring antimicrobial resistance Outcomes across Realms) is an R package designed to facilitate meta-analysis of AMR data across these domains.

This vignette provides an introduction to the basic functionality of the METEOR package, demonstrating how to import data, perform meta-analyses, and create visualizations.

## Installation

You can install the METEOR package from GitHub:

```{r, eval=FALSE}
# Install devtools if not already installed
if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}

# Install meteor from GitHub
devtools::install_github("username/meteor")
```

Once installed, you can load the package:

```{r, eval=FALSE}
library(meteor)
```

## Basic Workflow

### 1. Setting Up a Project

To get started, you can create a new METEOR project structure:

```{r, eval=FALSE}
initialize_meteor_project("my_amr_analysis")
```

This creates a directory structure with folders for data, analysis, results, and reports.

### 2. Importing Data

METEOR supports importing data from various file formats:

```{r, eval=FALSE}
# Import from CSV
amr_data <- import_amr_data("path/to/data.csv")

# Import from Excel
amr_data <- import_amr_data("path/to/data.xlsx", sheet = "Resistance_Data")

# Using built-in data
data(human_data)
```

### 3. Data Validation and Standardization

Before analysis, you should validate and standardize your data:

```{r, eval=FALSE}
# Validate data
validated_data <- validate_data(amr_data, domain = "human")

# Check data quality
quality_check <- check_data_quality(validated_data)
print(attr(quality_check, "quality"))

# Standardize data
std_data <- standardize_amr_data(validated_data, domain = "human")
```

### 4. Meta-Analysis

Perform meta-analysis on the standardized data:

```{r, eval=FALSE}
# Basic meta-analysis
meta_results <- calculate_pooled_rate(std_data)

# Meta-analysis by pathogen and antibiotic
grouped_meta <- calculate_pooled_rate(
  std_data,
  by = c("pathogen", "antibiotic"),
  method = "random"
)

# Analyze heterogeneity
heterogeneity <- analyze_heterogeneity(grouped_meta)
print(heterogeneity)

# Perform subgroup analysis
subgroup_results <- perform_subgroup_analysis(
  std_data,
  by = c("region"),
  meta_var = c("pathogen", "antibiotic")
)

# Sensitivity analysis
sensitivity <- perform_sensitivity_analysis(std_data)
```

### 5. Visualization

Create visualizations of the meta-analysis results:

```{r, eval=FALSE}
# Create forest plot
forest_plot <- create_forest_plot(grouped_meta)

# Create interactive forest plot
interactive_forest <- create_interactive_forest(grouped_meta)

# Create cumulative forest plot
cumulative_forest <- create_cumulative_forest(meta_results, order_by = "year")
```

### 6. Reporting

Generate reports of your findings:

```{r, eval=FALSE}
# Generate summary report (not implemented yet)
# generate_summary_report(meta_results, output_file = "amr_summary_report.html")
```

## Example: Meta-Analysis of E. coli Resistance

Below is a complete example workflow using sample data:

```{r, eval=FALSE}
# Load package
library(meteor)

# Load sample data
data(standardized_human_data)

# Filter for E. coli and ciprofloxacin
ecoli_cip <- standardized_human_data %>%
  dplyr::filter(
    pathogen == "Escherichia coli",
    antibiotic == "Ciprofloxacin"
  )

# Perform meta-analysis
ecoli_meta <- calculate_pooled_rate(ecoli_cip)

# Create forest plot
create_forest_plot(
  ecoli_meta,
  title = "Meta-Analysis of E. coli Resistance to Ciprofloxacin",
  xlab = "Resistance Rate"
)

# Analyze heterogeneity
heterogeneity <- analyze_heterogeneity(ecoli_meta, detailed = TRUE)
print(heterogeneity)

# Check for publication bias
bias <- assess_publication_bias(ecoli_cip)
```

## Using the Shiny App

For a more interactive experience, you can use the METEOR Shiny application:

```{r, eval=FALSE}
# Launch the METEOR Shiny app
launch_meteor()
```

This opens an interactive web application where you can import data, perform analyses, and create visualizations through a graphical user interface.

## Conclusion

METEOR provides a comprehensive set of tools for meta-analysis of antimicrobial resistance data. This vignette covered the basic workflow - importing data, performing meta-analyses, and creating visualizations. For more detailed information on specific functions, please refer to the function documentation.

## References

1. DerSimonian, R., & Laird, N. (1986). Meta-analysis in clinical trials. Controlled Clinical Trials, 7(3), 177-188.
2. Schwarzer, G. (2007). meta: An R package for meta-analysis. R News, 7(3), 40-45.
3. World Health Organization. (2014). Antimicrobial resistance: global report on surveillance. World Health Organization. 